---
title: "COMPSCIX 415.2 Homework 5/Midterm"
author: "Vishnu Vardhan"
date: "3/4/2018"
output: 
   pdf_document:
      toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('tidyverse')
library('dplyr')
library('ggthemes')
```


[Link to Github](https://github.com/vardhanv/compscix-415-2-assignments)

# 1. RStudio and R Markdown (3 points)

## Q.1
Use markdown headers in your document to clearly separate each midterm question and add a table of contents to your document.

#### Answer
Assignment submitted in the requested format.


# 2. The tidyverse packages (3 points)

## Q.1 
Can you name which package is associated with each task below?

- Plotting

- Data munging/wrangling

- Reshaping (speading and gathering) data

- Importing/exporting data



#### Answer

- Plotting - ggplot

- Data munging/wrangling - dplry

- Reshaping - tidyr, a part of the tidyverse universe

- Importing / exporting - This is spread accross the utils and the base packages


## Q.2 
Now can you name two functions that you’ve used from each package that you listed above for these tasks?

- Plotting

- Data munging/wrangling

- Reshaping data 

- Importing/exporting data (note that readRDS and saveRDS are base R functions) -

#### Answer

- Plotting - ggplot 
  + geom_point
  + geom_boxplot
  
- Data munging/wrangling - dplry
  + select
  + filter
  
- Reshaping - tidyr, a part of the tidyverse universe
  + spread
  + gather
  
  
- Importing / exporting - This is spread accross the utils and the base packages
  + read.csv
  + saveRDS
  


# 3. R Basics (1.5 points)

## Q.1
Fix this code with the fewest number of changes possible so it works:

My_data.name___is.too00ooLong! <- c( 1 , 2 , 3 )

#### Answer

```{r}
My_data.name___is.too00ooLong <- c( 1 , 2 , 3 )
My_data.name___is.too00ooLong
```

## Q.2 
Fix this code so it works:
my_string <- C('has', 'an', 'error', 'in', 'it)

#### Answer

```{r}
my_string <- c('has', 'an', 'error', 'in', 'it')
my_string
```

## Q.3
Look at the code below and comment on what happened to the values in the vector.

my_vector <- c(1, 2, '3', '4', 5)
my_vector

[1] "1" "2" "3" "4" "5"

#### Answer

Rephrasing the help in my own words.

c is a generic function that combines all its arguments into a vector or a list, coercing all the elements to a common type. 

The output type is determined from the highest type in the following hierarchy:

NULL < raw < logical < double < complex < character < list < expression

In this case, because character is larger than any of the numeric options, the entire vector is a vector of characters.


# 4. Data import/export (3 points)

## Q.1
Download the rail_trail.txt file from Canvas (in the Midterm Exam section here) and successfully import it into R. Prove that it was imported successfully by including your import code and taking a glimpse of the result.

#### Answer

```{r}
rail_trail <- read.delim('rail_trail.txt', header = TRUE, sep = "|")
glimpse(rail_trail)
```


## Q.2
Export the file into an R-specific format and name it “rail_trail.rds”. Make sure you define the path correctly so that you know where it gets saved. Then reload the file. Include your export and import code and take another glimpse.

#### Answer

```{r}
saveRDS(rail_trail, "rail_trail.rds")
new_rail_trail <- readRDS("rail_trail.rds")
glimpse(new_rail_trail) 
```


# 5. Visualization (6 points)

## Q.1
Critique this graphic: give only three examples of what is wrong with this graphic. Be concise.


#### Answer

1. The diagram uses 'area of a circle' to distinguish between the sizes of each segment. Bar graphs would be better, since they are single dimensional and easier to understand.

2. These are two seperate charts, but they look like one. The first chart is a chart with three ranges (<45, 45 to 64, and >64), the second chart is a men vs women chart. This simple difference is not easily visible with how it is layed out currently.

3. With the way the data is currently layed out it is not clear that yes/no data points are proportions. Though it says proportions in the title, it should visually be represented (and the numbers should show %)

## Q.2
Reproduce this graphic using the diamonds data set.

#### Answer

I was not able to get an exact replica, since it was not clear to me, what was used to re-order the boxplots. I used the lower of the inter-quartile range to get as close as i could.


```{r}
ggplot(data = diamonds) + 
  geom_boxplot(mapping = aes(x = reorder(cut, carat, function(x){quantile(x)[2] * -1}), 
                             y = carat, fill = color), 
               position = "identity") + 
  labs(x = "CUT OF DIAMOND", y = "CARAT OF DIAMOND") + 
  coord_flip()
```


## Q.3
The previous graphic is not very useful. We can make it much more useful by changing one thing about it. Make the change and plot it again.

#### Answer

I changed the position of the boxplots. So, that all the boxplots are visible, versus being overlapped.

```{r}
ggplot(data = diamonds) + 
  geom_boxplot(mapping = aes(x = reorder(cut, carat, function(x){quantile(x)[2] * -1}), 
                             y = carat, fill = color)) + 
  labs(x = "CUT OF DIAMOND", y = "CARAT OF DIAMOND") + 
  coord_flip() +
  theme_bw()
```


# 6. Data munging and wrangling (6 points)

## Q.1
Is this data “tidy”? If yes, leave it alone and go to the next problem. If no, make it tidy. Note: this data set is called table2 and is available in the tidyverse package. It should be ready for you to use after you’ve loaded the tidyverse package.

```{r}
table2
```

#### Answer

This data is not tidy.

Take the example of Afghanistan. For one observation in the year 1999, we have two rows, one for cases and one for population. To make this data tidy, there needs to be one observation per row, which we can achieve with a "spread".

```{r}
table2 %>% spread(type,count)
```


## Q.2
Create a new column in the diamonds data set called price_per_carat that shows the price of each diamond per carat (hint: divide). Only show me the code, not the output.

#### Answer

```{r, results='hide'}
diamonds %>% mutate(price_per_carat = price / carat)
```




## Q.3
For each cut of diamond in the diamonds data set, how many diamonds, and what proportion, have a price > 10000 and a carat < 1.5? There are several ways to get to an answer, but your solution must use the data wrangling verbs from the tidyverse in order to get credit.

• Do the results make sense? Why?

• Do we need to be wary of any of these numbers? Why?


#### Answer


```{r}
diamond_target <- diamonds %>% 
  mutate (target_segment = (price > 10000 & carat < 1.5)) %>% 
  group_by(cut) %>% 
  summarise( target_propotion = (sum(target_segment)/length(target_segment))*100, 
             target_count     = sum(target_segment))

ggplot(data = diamond_target, mapping = aes(x = cut, y = target_propotion)) + 
  geom_bar(stat = "identity") +
  geom_text(aes(y= target_propotion + 0.1,label = target_count)) +
  geom_text(aes(y= target_propotion/2,
                label = paste("paste(",round(target_propotion,2),",\" %\")")), 
            parse = TRUE, color = "green") +
  labs(title = "Diamonds priced > $10,000 and less than 1.5 carats", x = "Cut", 
       y = "% of Diamonds") +
  theme_bw()
```


As illustrated in the table above, there are 485 ideal diamonds, and they comprise 2.25% of all ideal diamonds. This makes sense, since as the diamon is more ideal, small diamonds are more expensive. Similarly, most fair diamonds won't have the same price as any of the others.

It is interesting that very-good and premium diamonds are the same. Which implies that we are missing some other parameter, likely clarity, colour or some such variable.

ps: I referred to this stack overflow thread to get the geom_text labels right.
https://stackoverflow.com/questions/4408414/cannot-concatenate-more-than-3-elements-in-an-expression-for-ggplot2s-geom-text



# 7. EDA (6 points)

Take a look at the txhousing data set that is included with the ggplot2 package and answer these questions:

## Q.1
During what time period is this data from?

#### Answer

The data is from Jan/2000 to July/2015

```{r}
txhousing %>% arrange(year,month)
txhousing %>% arrange(desc(year), desc(month))
```


## Q.2
How many cities are represented?

#### Answer

46 Cities are represented

```{r}
txhousing %>% select(city) %>% unique() %>% count()
```



## Q.3
Which city, month and year had the highest number of sales?

#### Answer

Houston, in July/2015 had sales volume of $ 2.568 B

```{r}
txhousing %>% arrange(desc(volume)) %>% top_n(1,volume)
```


## Q.4
What kind of relationship do you think exists between the number of listings and the number of sales? Check your assumption and show your work.


#### Answer

My assumption was the number of sales would increase as the listings increase, but clearly the law of diminishing returns apply.

```{r}
ggplot(data = txhousing,mapping = aes(x=listings, y = sales)) + 
  geom_point(alpha=0.4) + 
  geom_smooth() +
  theme_bw()
```

## Q.5
What proportion of sales is missing for each city?

#### Answer

Proportion of sales missing is:

```{r}
t <- txhousing %>% 
  mutate(valid_sales = !is.na(sales)) %>% 
  group_by(city) %>% 
  summarize(proportion = round(1 - sum(valid_sales)/length(valid_sales),2)) %>% 
  arrange(desc(proportion))
t
```

Only plotting cities with invalid sales.

```{r}
t <- t %>% filter(proportion > 0)

ggplot(data = t, 
       mapping = aes(x=reorder(city,proportion,min),
                     y=proportion))+ 
  geom_col()  + 
  geom_text(aes(label=round(proportion,2)), nudge_y = 0.02, size = 3) +
  labs(x = "Cities", y = "Proportion of Invalid Sales") +
  coord_flip() +
  theme_bw()
```


## Q.6
Looking at only the cities and months with greater than 500 sales:

- Are the distributions of the median sales price (column name median), when grouped by city, different? The same? Show your work.

- Any cities that stand out that you’d want to investigate further?

- Why might we want to filter out all cities and months with sales less than 500?


#### Answer

Cities with sales less than 500, are very small in value (1/8), but very large by volume (6/7), skewing our results, and it makes sense to eliminate them.

```{r}
txhousing %>% group_by(sales < 500) %>% summarise(sum(volume))
txhousing %>% group_by(sales < 500) %>% summarise(n())
```


Looking at the distribution shows that the median when grouped by cities is clearly different.

```{r}
tx_500 <- txhousing %>% filter(sales > 500)
ggplot(data = tx_500, mapping = aes(x=reorder(city,median,mean), y=median)) + 
  geom_boxplot() + 
  labs(x = "Cities", y = "Median Sale Price") + 
  coord_flip() +
  theme_bw()
```

Some of the interesting elements of the interesting cities are

1. Those were the median price is fairly high compared to the IQR
    - San Antonio
    - El Paso
    - Austin

2. Cities with lots of outliers may also beg further study
    - Denton County
    - Dallas
    - Houston
    



# 8. Git and Github (1.5 points)

## Q.1
To demonstrate your use of git and Github, at the top of your document put a hyperlink to your Github repository.

#### Answer

Git hub link added at the top of the document.

## Q.2
Once you are finished with your midterm, commit your final changes with the comment “finished the midterm- woohoo” and push your R Markdown file and your html or pdf file to Github.

#### Answer

done
